%% 
%% Copyright 2007-2020 Elsevier Ltd
%% 
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%% 

%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%% SP 2008/03/01
%%
%% 
%%
%% $Id: elsarticle-template-num.tex 190 2020-11-23 11:12:32Z rishi $
%%
%%
%\documentclass[preprint,12pt]{elsarticle}
\documentclass[final,3p,times,dvipsnames]{elsarticle}


%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm}
%% Our added packages
\usepackage{personal}

\journal{Journal of Computational Physics}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \affiliation{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}
%% \fntext[label3]{}

\title{Sampling-based sensitivity indices for stochastic solvers with application to Monte Carlo radiation transport}

%% use optional labels to link authors explicitly to addresses:
% \author[label1,label2]{}
% \affiliation[label1]{organization={},
%             addressline={},
%             city={},
%             postcode={},
%             state={},
%             country={}}
       
\author[1]{Kayla~B. Clements\corref{cor1}}
%\cormark[1]
%\fnmark[thanks]
\ead{clemekay@oregonstate.edu,kbcleme@sandia.gov}
%\ead[url]{}
%\credit{test}       
       
\author[2]{Gianluca Geraci\corref{cor1}}
%\cormark[1]
%\fnmark[thanks]
\ead{ggeraci@sandia.gov}
%\ead[url]{}
%\credit{test}     

\author[2]{Aaron~J. Olson}
%\cormark[1]
%\fnmark[thanks]
\ead{aolson@sandia.gov}
%\ead[url]{}
%\credit{test}

\author[1]{Todd~S. Palmer}
%\cormark[1]
%\fnmark[thanks]
\ead{todd.palmer@oregonstate.edu}
%\ead[url]{}
%\credit{test} 

\cortext[cor1]{Corresponding authors}
            
% Address/affiliation
\affiliation[1]{organization={Oregon State University},
            addressline={Address}, 
            city={Corvallis},
%           citysep={}, % Uncomment if no comma needed between city and postcode
            % postcode={XXX}, 
            state={OR},
            country={USA}}           
            

\affiliation[2]{organization={Sandia National Laboratories},
            addressline={P.O. Box 5800, Mail Stop 1318}, 
            city={Albuquerque},
%           citysep={}, % Uncomment if no comma needed between city and postcode
            postcode={87185-1318}, 
            state={NM},
            country={USA}}           


\begin{abstract}
In computational modeling, global sensitivity analysis aims to characterize how input variability affects output variability.
Sobol' indices, a variance-based tool for global sensitivity analysis, rank parameters in order of importance to model response across the entire combined input parameter space.
Accurate and efficient methods for computing Sobol' indices have been widely researched for deterministic simulators, in which multiple evaluations of the same input will produce identical outputs.
Stochastic simulators, on the hand, have an intrinsic randomness and produce different outputs for multiple evaluations of the same input. 
This introduces additional variability to model output, complicating the use of traditional methods for computing Sobol' indices. 
In this paper, we focus on computing Sobol' indices that are unbiased by solver noise without needing to over-resolve each evaluation of the stochastic simulator.
We propose doing so using variance deconvolution, in which we explicitly calculate the variance due to the solver and remove it from the total observed variance.
The proposed method is applied to two examples: the Ishigami function that is commonly used as a test case for Sobol' indices and a neutron-transport case study.
The results confirm the convergence of the approach and highlight the approach's utility particularly when the indices are not near-zero and when there is a large amount of solver noise.
\end{abstract}


%%Research highlights
\begin{highlights}
\item An approach is developed for global sensitivity analysis using stochastic computational solvers.
\item Variance deconvolution, in which solver variance is explicitly calculated and removed from the total observed variance, is briefly reviewed.
\item Variance deconvolution (var-d) is combined with existing methods for Sobol' indices.
\item Var-d Sobol' indices are shown to converge to the true solution for an analytic test case.
\item Monte Carlo radiation transport is used for an additional numerical example.
\end{highlights}


\begin{keyword}
global sensitivity analysis \sep Monte Carlo radiation transport \sep stochastic solvers \sep sobol indices
\end{keyword}


\end{frontmatter}


\section{Introduction}
\label{sec:intro}
\input{section-intro.tex}


\section{Background and theory on ANOVA}
\label{sec:anova}
\input{section-anova.tex}


\section{Sampling-based estimators for sensitivity indices}
\label{sec:sampling}
\input{section-sampling.tex}


\section{Introduction to variance deconvolution}
\label{sec:deconvolution}
\input{section-stochastic.tex}


\section{GSA with variance deconvolution}
\label{sec:gsa-deconvolution}
\input{section-stochastic-gsa.tex}


\section{Numerical results}
\label{sec:results}
\input{section-results.tex}


\section{Conclusion}
\label{sec:conclusion}
\input{section-conclusions.tex}


\section{Mean-squared error from asymptotic limits}
\label{appx:mse}
\input{section-appx-mse.tex}


\bibliographystyle{elsarticle-num}
\bibliography{biblio}


\end{document}
\endinput

